---
title: "GUstica_Report.html"
author: "De Vivo Massimo"
date: "2023-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("C:\\Users\\Max\\Downloads")
load("Agrimonia_stations.RData")
ls()
dim(table(agri$NameStation))
head(agri)
dim(agri)
names(agri)

```
Datas are loaded into agri's table.

# Variabile risposta y
PM10: Particulate matter with an aerodynamic diameter of less than 10 μm


We chose PM10 as the response variable because it fits our model best.

# Explanatory variable x to answer the research question
LI_bovine: Municipal density of bovines related to AQ stations

We chose the density of bovine farms as the explanatory variable because we are interested in understanding how livestock farming affects air quality.

Additionally, we used the variables
WE_temp_2m: Temperature of air at 2 m above the surface of land, sea, or inland water
WE_wind_speed_10_mean: Mean wind speed intensity at a height of 10 m above the Earth's surface

To provide a more realistic comparison with air pollution, as the bovine density variable alone did not yield consistent results.

# Research Question
What impact do bovine farms have on air quality? 
What effect does increasing or decreasing their number have on the PM10 pollutant?

# Approach
We select the relevant days

```{r}
agri$NameStation = as.factor(agri$NameStation)
agri$Province = as.factor(agri$Province)
summary(agri$NameStation)

agri_da = agri[agri$Time=="2017-07-11",]
agri_da2= agri[agri$Time=="2017-01-11",]
dim(agri_da)

```
Abbiamo selezionato un giorno estivo e uno invernale per avere il confronto tra due stagioni opposte.

```{r}
#estate
agri_da_y = agri_da[!is.na(agri_da$LI_bovine),]
dim(agri_da_y)
#inverno
agri_da_x = agri_da2[!is.na(agri_da$LI_bovine),]
dim(agri_da_x)
```
Abbiamo rimosso dalla tabella tutte le stazioni che hanno la variabile degli allevamenti di bovini con valore mancante.

```{r}
cat("variabilità bovini estate", "\n")
summary(agri_da_y$LI_bovine)
cat("\n", "variabilità bovini inverno", "\n")
summary(agri_da_x$LI_bovine)
```
Abbiamo stampato la variabilità del numero di bovini in estate e in inverno per verificare la differenza nei due casi.


```{r}

# ESTATE
attach(agri_da_y)
plot(agri_da_y$NameStation, AQ_pm10, xlab = "" , ylab = "pm 10 estate", las = 2, cex.axis = 0.4)
legend("topright", legend = "pm 10")

plot(agri_da_y$NameStation, AQ_pm25, xlab = "" , ylab = "pm 2.5 estate", las = 2, cex.axis = 0.4)
legend("topright", legend = "pm 2.5")

plot(agri_da_y$NameStation, AQ_nox, xlab = "" , ylab = "nox estate", las = 2, cex.axis = 0.4)
legend("topright", legend = "nox")

#INVERNO
attach(agri_da_x)
plot(agri_da_x$NameStation, AQ_pm10, xlab = "" , ylab = "pm 10 inverno", las = 2, cex.axis = 0.4)
legend("topright", legend = "pm10")

plot(agri_da_x$NameStation, AQ_pm25, xlab = "" , ylab = "pm 2.5 inverno", las = 2, cex.axis = 0.4)
legend("topright", legend = "pm 2.5")

plot(agri_da_x$NameStation, AQ_nox, xlab = "" , ylab = "nox inverno", las = 2, cex.axis = 0.4)
legend("topright", legend = "nox")

```
Abbiamo creato i grafici con la distribuzione delle diverse variabili che rappresentano i fattori inquinanti dell'aria per stazione nel singolo giorno, separando estate ed inverno. 

Da questi possiamo notare che c'è differenza di inquinamento tra estate ed inverno.

```{r}
mean_pm10_estate <- mean(agri_da_y$AQ_pm10, na.rm = TRUE)
mean_pm10_inverno <- mean(agri_da_x$AQ_pm10, na.rm = TRUE)

mean_pm25_estate <- mean(agri_da_y$AQ_pm25, na.rm = TRUE)
mean_pm25_inverno <- mean(agri_da_x$AQ_pm25, na.rm = TRUE)

mean_nox_estate <- mean(agri_da_y$AQ_nox, na.rm = TRUE)
mean_nox_inverno <- mean(agri_da_x$AQ_nox, na.rm = TRUE)

cat("Media PM10 in estate:", mean_pm10_estate, "\n")
cat("Media PM10 in inverno:", mean_pm10_inverno, "\n")

cat("Media PM2.5 in estate:", mean_pm25_estate, "\n")
cat("Media PM2.5 in inverno:", mean_pm25_inverno, "\n")

cat("Media NOx in estate:", mean_nox_estate, "\n")
cat("Media NOx in inverno:", mean_nox_inverno, "\n")
```
Per avere ulteriori conferme, abbiamo calcolato anche la media dei valori dei vari inquinanti e abbiamo notato che in inverno l'inquinamento è più alto.


# Analisi variabile risposta

```{r}
#ESTATE pm10
# histogram
hist(agri_da_y$AQ_pm10, main = "Distribuzione di pm10 in estate", xlab = "pm10", col = "skyblue", border = "black")

#plot density
plot(density(agri_da_y$AQ_pm10, na.rm = TRUE), main = "Densità di pm10", xlab = "pm10", col = "skyblue", lwd = 2)

plot(density(log(agri_da_y$AQ_pm10),na.rm = TRUE), main = "Densità di pm10 log", xlab = "pm10", col = "purple", lwd = 2)

plot(density(sqrt(agri_da_y$AQ_pm10), na.rm = TRUE), main = "Densità di pm10 radice quadrata", xlab = "pm10", col = "blue", lwd = 2)


plot(density((agri_da_y$AQ_pm10)^(1/3), na.rm = TRUE), main = "Densità di pm10 radice cubica", xlab = "pm10", col = "red", lwd = 2)


# INVERNO pm10

# histogram
hist(agri_da_x$AQ_pm10, main = "Distribuzione di pm10 in inverno", xlab = "pm10", col = "skyblue", border = "black")

#plot density
plot(density(agri_da_x$AQ_pm10, na.rm = TRUE), main = "Densità di pm10", xlab = "pm10", col = "skyblue", lwd = 2)

plot(density(log(agri_da_x$AQ_pm10),na.rm = TRUE), main = "Densità di pm10 log", xlab = "pm10", col = "purple", lwd = 2)

plot(density(sqrt(agri_da_x$AQ_pm10), na.rm = TRUE), main = "Densità di pm10 radice quadrata", xlab = "pm10", col = "blue", lwd = 2)


plot(density((agri_da_x$AQ_pm10)^(1/3), na.rm = TRUE), main = "Densità di pm10 radice cubica", xlab = "pm10", col = "red", lwd = 2)

```
Abbiamo fatto diversi grafici per normalizzare la distribuzione di pm10.
Per l'estate abbiamo scelto la radice cubica, mentre per l'inverno il grafico senza approssimazioni.


# Correlazione
```{r}
#ESTATE
library(corrplot)
cor_matrix_estate <- cor(agri_da_y[, c("AQ_pm10", "AQ_pm25", "LI_bovine", "AQ_nox")],use = "complete.obs")
cat("Correlazione variabili in estate")
cat("\n")
print(cor_matrix_estate)
cat("\n")
corrplot(cor_matrix_estate, method = "color", type = "upper", order = "hclust", tl.cex = 0.7)

#INVERNO
cor_matrix_inverno <- cor(agri_da_x[, c("AQ_pm10", "AQ_pm25", "LI_bovine", "AQ_nox")],use = "complete.obs")
cat("Correlazione variabili in inverno")
cat("\n")
print(cor_matrix_inverno)
corrplot(cor_matrix_inverno, method = "color", type = "upper", order = "hclust", tl.cex = 0.7)

```
Abbiamo calcolato la correlazione dei vari inquinanti in estate e in inverno con la variabile bovini.
Possiamo notare che la correlazione tra bovini ed inquinanti:
- in estate è positiva ma debole, di conseguenza possiamo dire che le emissioni dei bovini hanno un'influenza minima sulla qualità dell'aria.
- in inverno è debolmente negativa, ovvero le emissioni di bovini non influiscono sull'inquinamento. Una spiegazione potrebbe essere che, essendoci molti altri agenti inquinanti, l'impatto dei bovini è irrilevante. 


# Modello lineare 

```{r}
# estate
model_pm10 <- lm((AQ_pm10)^(1/3) ~ LI_bovine, data = agri_da_y)
model_pm10_2 <- lm((AQ_pm10)^(1/3) ~ LI_bovine + WE_temp_2m +WE_wind_speed_10m_mean,  data = agri_da_y)
cat("Modello lineare a una variabile (estate)", "\n")
summary(model_pm10)
cat("Modello lineare a più variabili (estate)", "\n")
summary(model_pm10_2)
```
In the single-variable linear model, we observe that an increase of one unit in LI_bovine is associated with an estimated increase of 0.00177 in the cubic mean of PM10.

The p-value associated with LI_bovine is 0.00697, which is less than 0.05, indicating that the coefficient is statistically significant.

The model has an R-squared of 0.07415, meaning that 7.4% of the variation in the cubic mean of AQ_PM10 can be explained by the LI_bovine variable.

In the multi-variable linear model, we observe the following:

- LI_bovine: The coefficient is 0.0009026, and the associated p-value is 0.104578. It is not statistically significant at a 5% significance level.

- WE_temp_2m: The coefficient is 0.0371794, and the associated p-value is very low (0.000489). This indicates that WE_temp_2m is statistically significant and positively associated with the cubic mean of PM10.

- WE_wind_speed_10m_mean: The coefficient is 0.0463938, but the p-value is 0.399096, indicating that this coefficient is not statistically significant at a 5% significance level.

The model has an R-squared of 0.3763, meaning that 37.63% of the variation in the cubic mean of AQ_PM10 can be explained by the predictive variables in the model.

In conclusion, we can state that the second multi-variable model seems to explain a greater percentage of the variation in the cubic mean of PM10 compared to the first single-variable model.

```{r}
#inverno

model_pm10_inv <- lm(AQ_pm10 ~ LI_bovine, data = agri_da_x)
model_pm10_inv_2 <- lm(AQ_pm10 ~ LI_bovine + WE_temp_2m + WE_wind_speed_10m_mean, data = agri_da_x)
cat("Modello lineare a una variabile (inverno)", "\n")
summary(model_pm10_inv)
cat("Modello lineare a più variabili (inverno)", "\n")
summary(model_pm10_inv_2)

```

```{r}

par(mfrow = c(2, 2))

#ESTATE
plot(model_pm10)
plot(model_pm10_2)

#INVERNO
plot(model_pm10_inv)
plot(model_pm10_inv_2)

```
  
 


PREVISIONE ESTATE

```{r}


model_pm10_2_basso <- lm((AQ_pm10)^(1/3) ~ LI_bovine + WE_temp_2m +WE_wind_speed_10m_mean,  data = agri_da_y)
bovini=10
temp=agri_da_y$WE_temp_2m
wind=agri_da_y$WE_wind_speed_10m_mean

newdata<-data.frame(LI_bovine=bovini, WE_temp_2m=temp,WE_wind_speed_10m_mean=wind)
predict(model_pm10_2_basso,newdata,interval="predict", level=0.95)



model_pm10_2_alto <- lm((AQ_pm10)^(1/3) ~ LI_bovine + WE_temp_2m +WE_wind_speed_10m_mean,  data = agri_da_y)
bovini=10000
temp=agri_da_y$WE_temp_2m
wind=agri_da_y$WE_wind_speed_10m_mean

newdata<-data.frame(LI_bovine=bovini, WE_temp_2m=temp,WE_wind_speed_10m_mean=wind)
predict(model_pm10_2_alto,newdata,interval="predict", level=0.95)
```

PREVISIONE INVERNO
```{r}

model_pm10_inv_2_basso <- lm(AQ_pm10 ~ LI_bovine + WE_temp_2m +WE_wind_speed_10m_mean,  data = agri_da_x)
bovini=10
temp=agri_da_x$WE_temp_2m
wind=agri_da_x$WE_wind_speed_10m_mean

newdata<-data.frame(LI_bovine=bovini, WE_temp_2m=temp,WE_wind_speed_10m_mean=wind)
predict(model_pm10_inv_2_basso,newdata,interval="predict", level=0.95)



model_pm10_inv_2_alto <- lm(AQ_pm10 ~ LI_bovine + WE_temp_2m +WE_wind_speed_10m_mean,  data = agri_da_x)
bovini=10000
temp=agri_da_x$WE_temp_2m
wind=agri_da_x$WE_wind_speed_10m_mean

newdata<-data.frame(LI_bovine=bovini, WE_temp_2m=temp,WE_wind_speed_10m_mean=wind)
predict(model_pm10_inv_2_alto,newdata,interval="predict", level=0.95)

```


```